WSClean: an implementation of a fast, generic wide-field imager for radio astronomy

### 2014-04-01 articles/2014/Article-WSClean/wsclean.tex
Version as submitted\!

\ifpdf\pdfinfo{/Title      (WSClean: an implementation of a fast, generic wide-field imager)
               /Author     (A. R. Offringa et al.)
               /Keywords   (instrumentation: interferometers;methods: observational;techniques: interferometric;radio continuum: general)
        }
\else\usepackage{graphics}\fi

@@

%It is important to point out that all these measurement were performed by imaging a single polarisation. Making full-polarisation images with \textsc{wsclean} currently requires four runs. Imaging four Stokes parameters with \textsc{casa} costs slightly less than four times the time of imaging a single polarisation. With the settings of Table~\ref{tbl:default-parameters}, the difference is a factor of $3.5$, and imaging XX and YY together costs $1.9$ times imaging just XX.

@@

\begin{table*}
\caption{Configurations for which the computational cost is predicted. Columns: $\lambda$=wavelength; FOV=field of view; Beams=number of beams; Ant=number of elements; Res=angular resolution; $D$=maximum baseline; $\max \Delta z$=maximum differential elevation; $\Delta t$=correlator dump time; BW=bandwidth; and $\Delta \nu$=correlator frequency resolution.} \label{tbl:computational-cost-configurations}
\begin{tabular}{l|rrrrrrrrrr}
 Configuration&$\lambda$& FOV  & Beams & Ant & Res & $D$ & $\max \Delta z$ & $\Delta t$ & BW & $\Delta \nu$ \\
          & (m)       & (FWHM)&       &     &     & (km)                       & (m)             & (s)    & (MHz) &  (kHz) \\
\hline
 GLEAM    &  2   & 24.7\degree& 1    & 128 & 2'   & 2.9 & 5 & 2 & 32 & 40 \\
 EMU      & 0.2  & 1\degree   & 30   &  36 & 10''  & 6 & 0.2 &10 & 300 & 20\\
 MSSS low &  5   & 9.8\degree & 5    &  20 & 100'' & 5 & 2  &10 & 16 & 16 \\
 MSSS high&  2   & 3.8\degree & 5    &  40 & 120'' & 5 & 2  &10 & 16 & 16 \\
 LOFAR LBA NL& 5 & 4.9\degree & 1    &  38 &   3'' &180& 20 & 1 & 96 & 1 \\
 AARTFAAC &  5   & 45\degree  & 1    & 288 &  20'  & 0.3& 0.5 & 1 &  7 & 24 \\
 VLSS     & 4.1  & 14\degree  & 1    &  27 &  80'' &11.1& 0.2 &10 & 1.56 & 12.2 \\
 MeerKAT  & 0.2  &  1\degree  & 1    &  64 &   6'' &  8 & 1   & 0.5 & 750  & 50 \\
% SKA1 AA inner&2 &  5\degree  & 1    & 433 &  15'' &0.6& 0.5 &  10 & 250 & 1 \\
 SKA1 AA core&2  &  5\degree  & 1    & 866 &    3' &  3& 5 & 10 & 250 & 1 \\
 SKA1 AA full&2  &  5\degree  & 1    & 911 &   5'' &100& 50 & 0.6& 250 & 1 \\
\end{tabular}
\end{table*}

\begin{table*}
\caption{Predicted computational costs for configurations listed in Table~\ref{tbl:computational-cost-configurations}, based on multi-frequency synthesis with five major iterations observing for 1~h at ZA=20\degree with 1 polarisation. Predictions are for $w$-projection with \textsc{casa}; $w$-stacking with \textsc{wsclean}; $w$-stacked snapshots with \textsc{wsclean} using optimal snapshot duration; and a hybrid between $w$-projection and $w$-stacking using \textsc{casa} with optimal $\Delta w$. All have approximately equal accuracy.} \label{tbl:computational-cost-predictions}
\begin{tabular}{l|rrrr|rr}
               & \multicolumn{4}{c|}{Predicted computing cost on test computer} & min & best \\
 Configuration & $w$-projection & $w$-stacking & $w$-snapshot & hybrid & computing & FLOPS/float \\
 \hline
 GLEAM    &  65h 25m & 8h 04m &  8h 03m & 14h 10m   & 1.1 TFLOPS & $6.8 \times 10^2$    \\
 EMU      &    5.2 d &  2.9 d &   2.9 d & 4.4 d     & 9.7 TFLOPS & $2.1 \times 10^4$      \\
 MSSS low &  2h 16m  & 0h 57m &  0h 57m & 2h 10m    & 130 GFLOPS & $3.4 \times 10^3$   \\
 MSSS high&  7h 16m  & 3h 52m &  3h 52m & 5h 44m    & 530 GFLOPS & $3.4 \times 10^3$   \\
 LOFAR NL &  50.2 d  &  7.3 d &   7.2 d & 12.9 d    &  24 TFLOPS & $7.2 \times 10^2$ \\
 AARTFAAC &  2.5 d   &  1.2 d &   1.2 d &   2.5 d   & 4.1 TFLOPS & $6.8 \times 10^2$   \\
 VLSS     &  0h 09m  & 0h 01m &  0h 01m &  0h 08m   & 2.2 GFLOPS & $1.1 \times 10^3$     \\
 MeerKAT  &   10.8 d &  6.2 d &   6.2 d &  10.8 d   &  21 TFLOPS & $6.8 \times 10^2$ \\
% SKA1 AA inner&392 d &  228 d &   228 d &   392 d & & $6.8 \times 10^2$   \\
 SKA1 AA core&1581 d &  911 d &   911 d &  1570.8   & 3.0 PFLOPS & $6.8 \times 10^2$ \\
 SKA1 AA full& 643 yr& 48.8 yr & 48.8 yr& 84.1 yr   &  59 PFLOPS & $6.8 \times 10^2$ \\
\end{tabular}
\end{table*}

>>>>

\ifpdf\pdfinfo{/Title      (WSClean: an implementation of a fast, generic wide-field imager for radio astronomy)
               /Author     (A. R. Offringa et al.)
               /Keywords   (instrumentation: interferometers;methods: observational;techniques: interferometric;radio continuum: general)
        }
\else\usepackage{graphics}\fi

@@

@@

\begin{table*}
\caption{Configurations for which the computational cost is predicted. Columns: $\lambda$=wavelength; FOV=field of view; Beams=number of beams; Ant=number of elements; Res=angular resolution; $D$=maximum baseline; $\max \Delta z$=maximum differential elevation; $\Delta t$=correlator dump time; BW=bandwidth; and $\Delta \nu$=correlator frequency resolution.} \label{tbl:computational-cost-configurations}
\begin{tabular}{l|rrrrrrrrrr}
 Configuration&$\lambda$& FOV  & Beams & Ant & Res & $D$ & $\max \Delta z$ & $\Delta t$ & BW & $\Delta \nu$ \\
          & (m)       & (FWHM)&       &     &     & (km)                       & (m)             & (s)    & (MHz) &  (kHz) \\
\hline
 GLEAM    &  2   & 24.7\degree& 1    & 128 & 2'   & 2.9 & 5 & 2 & 32 & 40 \\
 EMU      & 0.2  & 1\degree   & 30   &  36 & 10''  & 6 & 0.2 &10 & 300 & 20\\
 MSSS low &  5   & 9.8\degree & 5    &  20 & 100'' & 5 & 2  &10 & 16 & 16 \\
 MSSS high&  2   & 3.8\degree & 5    &  40 & 120'' & 5 & 2  &10 & 16 & 16 \\
 LOFAR LBA NL& 5 & 4.9\degree & 1    &  38 &   3'' &180& 20 & 1 & 96 & 1 \\
 AARTFAAC &  5   & 45\degree  & 1    & 288 &  20'  & 0.3& 0.5 & 1 &  7 & 24 \\
 VLSS     & 4.1  & 14\degree  & 1    &  27 &  80'' &11.1& 0.2 &10 & 1.56 & 12.2 \\
 MeerKAT  & 0.2  &  1\degree  & 1    &  64 &   6'' &  8 & 1   & 0.5 & 750  & 50 \\
 SKA1 AA core&2  &  5\degree  & 1    & 866 &    3' &  3& 5 & 10 & 250 & 1 \\
 SKA1 AA full&2  &  5\degree  & 1    & 911 &   5'' &100& 50 & 0.6& 250 & 1 \\
\end{tabular}
\end{table*}

\begin{table*}
\caption{Predicted computational costs for configurations listed in Table~\ref{tbl:computational-cost-configurations}, based on multi-frequency synthesis with five major iterations observing for 1~h at ZA=20\degree with 1 polarisation. Predictions are for $w$-projection with \textsc{casa}; $w$-stacking with \textsc{wsclean}; $w$-stacked snapshots with \textsc{wsclean} using optimal snapshot duration; and a hybrid between $w$-projection and $w$-stacking using \textsc{casa} with optimal $\Delta w$. All have approximately equal accuracy.} \label{tbl:computational-cost-predictions}
\begin{tabular}{l|rrrr|rr}
               & \multicolumn{4}{c|}{Predicted computing cost on test computer} & min & best \\
 Configuration & $w$-projection & $w$-stacking & $w$-snapshot & hybrid & computing & FLOPS/float \\
 \hline
 GLEAM    &  65h 25m & 8h 04m &  8h 03m & 14h 10m   & 1.1 TFLOPS & $6.8 \times 10^2$    \\
 EMU      &    5.2 d &  2.9 d &   2.9 d & 4.4 d     & 9.7 TFLOPS & $2.1 \times 10^4$      \\
 MSSS low &  2h 16m  & 0h 57m &  0h 57m & 2h 10m    & 130 GFLOPS & $3.4 \times 10^3$   \\
 MSSS high&  7h 16m  & 3h 52m &  3h 52m & 5h 44m    & 530 GFLOPS & $3.4 \times 10^3$   \\
 LOFAR NL &  50.2 d  &  7.3 d &   7.2 d & 12.9 d    &  24 TFLOPS & $7.2 \times 10^2$ \\
 AARTFAAC &  2.5 d   &  1.2 d &   1.2 d &   2.5 d   & 4.1 TFLOPS & $6.8 \times 10^2$   \\
 VLSS     &  0h 09m  & 0h 01m &  0h 01m &  0h 08m   & 2.2 GFLOPS & $1.1 \times 10^3$     \\
 MeerKAT  &   10.8 d &  6.2 d &   6.2 d &  10.8 d   &  21 TFLOPS & $6.8 \times 10^2$ \\
 SKA1 AA core&1581 d &  911 d &   911 d &  1570.8   & 3.0 PFLOPS & $6.8 \times 10^2$ \\
 SKA1 AA full& 643 yr& 48.8 yr & 48.8 yr& 84.1 yr   &  59 PFLOPS & $6.8 \times 10^2$ \\
\end{tabular}
\end{table*}
